Seminar:\\ Quantitative Methoden

### 2024-02-15 draft2.tex
update morning

\documentclass[12pt,draft,a4paper]{article}
\usepackage[authoryear]{natbib}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{caption}

\usepackage{newtx}
\usepackage[a4paper, left=3cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage[noabbrev]{cleveref}
\captionsetup[table]{labelfont=bf,
                     labelsep=newline,
                     singlelinecheck=false}

@@

    \author{Moritz Trinkner}
    \title{Seminar:\\ Quantitative Methoden}
    \date{February 2024}
\end{titlepage}
\maketitle
\thispagestyle{empty}
\newpage
\tableofcontents
\thispagestyle{empty}
\newpage
\setcounter{page}{1}

@@

Table 1 shows the descriptive statistics for all the variables we used in our analysis. We have 40,295 mothers in our sample.
Noticeable values in the table are that only 13.5\% of mothers are working and accordingly the average hours worked per mother is only 5.
The maximum working hours might seem high,
but since only 1\% of employed mothers work more than 70 hours and there are reports of sweatshops in Jordan with up to 100h worked per week (Greenhouse and Barbaro, 2006) we deemed it plausible.
It is worth to note here, that the working hours per week, only looking at working mothers, is 40.2.
This suggests that part time work is almost non-existent in Jordan, in our sample less than 4\% of working mothers work less that 30 hours per week.
This lack of part time work can partially be of an explanation of the low workforce participation of women. 
Additionally, 97.6\% of mothers are married and 81.1\% have more than 1 child.
All of this points to the fact that Jordan is a highly conservative and religious country with strict gender roles.
In our sample 51.7\% of first-born kids are boys, being close to the natural sex ratio. Therefore the key identification assumption that gender of first born boy is random holds.

\begin{table}[ht]
    \caption{T-tests of mothers' characteristics on sex of the first-born child}
    \begin{tabularx}{\textwidth}{lllll}
        \hline
          & \begin{tabular}[c]{@{}l@{}}Mothers with\\ first-born boys\\ (St. err.)\end{tabular} & 
          \begin{tabular}[c]{@{}l@{}}Mothers with\\ first-born girls\\ (St. err.)\end{tabular} & 
          \begin{tabular}[c]{@{}l@{}}Difference\\ (St. err.) \\ \end{tabular} & 
          \begin{tabular}[c]{@{}l@{}}T-test statistic\\ (P-value) \\ \end{tabular} \\ \hline
    Employed              & 0.134           & 0.132        & -0.002   & -0.612       \\
                          & (0.002)         & (0.002)      & (0.003)  & (0.541)      \\
    Hours worked per week & 5.471           & 5.417        & -0.053   & -0.371       \\
                          & (0.100)         & (0.103)      & (0.143)  & (0.710)      \\
    Age                   & 30.643          & 30.721       & 0.078    & 1.344        \\
                          & (0.040)         & (0.042)      & (0.058)  & (0.179)      \\
    Married               & 0.975           & 0.976        & 0.001    & 0.833        \\
                          & (0.001)         & (0.001)      & (0.002)  & (0.405)      \\
    More than 1 child     & 0.805           & 0.819        & 0.014    & 3.617        \\
                          & (0.003)         & (0.003)      & (0.004)  & (0.000)      \\
                          \hline
                        \end{tabularx}

@@

\begin{table}[ht]% not h om its own
    \caption{First child gender, fertility and marital status of the mother}

@@

    \hline & \\[-1.0em]
    Hours worked per week & & Probability of working &  \\
    \hline & \\[-1.0em]
        sFirst-born boy          & 0.086           & First-born boy           & 0.003           \\
        St. err.                & 0.141           & St. err.                 & 0.003           \\
        \\[-1.0em]
        Baseline Girl:          & 5.378           & Baseline Girl:           & 0.133           \\
        St. err.                & 1.524           & St. err.                 & 0.036           \\
        Percent effect          & 1.596           & Percent effect           & 2.172           \\
        No. obs.                & 40,295          & No. obs.                 & 40,295          \\
    \hline
    \end{tabularx}

@@

\begin{table}[ht]% not h om its own
    \caption{First child gender, fertility and marital status of the mother}

@@

The following Section all summarizes the paper “Did the minimum wage reduce the gender wage gap in Germany?” by \citet{CALIENDO22}. It addresses the question if the minimum wage can have an effect on the gender wage gap. The reason why this effect might appear is that women are in most countries and regions disproportionally represented among low paid workers. \citep{kahn2015wage}
This means that an increase in the minimum wage would affect women's average wages more than mens, leading to a decrease in the gender wage gap.
The minimum wage is a national wage floor, that is binding for almost all employees. Germany is a special research case, since it was one of the last OECD countries to introduce a Minimum wage. When it was introduced in 2015 at 8.50€ it was one of the highest in the EU in terms of purchasing power parity. It affected 10\% of the workforce, and 2/3 of those affected low income workers were women.  This strong increase makes Germany a great natural experiment to observe effects of the minimum wage, in this case the effect of the gender wage gap. 
%Short preview of results in introduction
%Detailed OECD Data from 2014 First Decile comparison to average
% not a panel study

@@

% Maybe explain ex
The Method used in the paper is a difference-in-difference approach with fixed effects. Since there are no legislative differences for the federal wage in difference regions in Germany, the authors use regional differences in Germany to assign Germany to the control or treatment group.
%Why fraction of in relation to women and not the whole workforce
%Not dependent on gender wage gap neither on fraction of men working under the minimum wage
%"We generate the gender wage gap at the regional level as the diﬀerence between men’s and women’s wages divided by men’s wages."
This is followed after \citet{Card1992} who did the same in an analysis of the minimum wage in the United States. 
They  use labor market regions to split the country in 257 different regions. 
The fraction of women that worked below the minimum wage in 2014 in relation to all employed women is used to see how 
strong %better word 
the introduction of the minimum wage affects the employed women in the region. 
\Citeauthor{CALIENDO22} call this the bite, which can be visualized by how deep the minimum wage bites into the wage distribution.
%median bite level of 17.15%.
The fraction for all regions is used to separate them at the median, so that regions with more than 17.15\% of women affected by the minimum wage are in the treatment group and regions below the control group.

@@

\subsection{Main Result}
% • What is the main result? And what is the implication of this finding?
% Summarize and synthesize the results, provide relevant background information
\begin{table}[btp] %Main Results
    \caption{Fixed effects regressions of wage gaps at 10th percentile, 25th percentile and mean (in labour market regions).}
    \centering
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{lllllll}
    \hline & \\[-1.0em]
                              & (1)       & (2)       & (3)       & (4)       & (5)       & (6)       \\ 
    \hline & \\[-1.0em]
    \multicolumn{7}{c}{A: Wage Gap at p10}                                                            \\
    \hline & \\[-1.0em]
    Bite x 2018               & -5.228*** & -5.085*** & -4.870*** & -4.735*** & -4.550*** & -4.550*** \\
    2018                      & -3.568*** & -4.286*** & -4.539*** & -5.419*** & -3.684    & -3.754    \\
    GDP per capita ($t$ - 1)   &           & 0.153     & 0.152     & 0.126     & 0.147     & 0.158     \\
    Pop. Density ($t$ - 1)     &           &           & 0.024     & 0.020     & 0.017     & 0.015     \\
    Share of Women ($t$ - 1)   &           &           &           & -2.722    & -2.837    & -2.610    \\
    Empl. Rate Women ($t$ - 1) &           &           &           &           & -0.412    & -0.457    \\
    Childcare 0-2 ($t$ - 1)    &           &           &           &           &           & 0.044     \\
    Childcare 3-5 ($t$ - 1)    &           &           &           &           &           & -0.061    \\ 
    \hline & \\[-1.0em]
    \multicolumn{7}{c}{B: Wage Gap at p25}                                                            \\ 
    \hline & \\[-1.0em]
    Bite x 2018               & -3.352*** & -2.814*** & -3.122*** & -3.111*** & -3.191*** & -3.336*** \\
    2018                      & -1.454**  & -4.154*** & -3.792*** & -3.865**  & -4.612    & -4.079    \\
    GDP per capita ($t$ - 1)   &           & 0.576**   & 0.578**   & 0.576**   & 0.567**   & 0.541**   \\
    Pop. Density ($t$ - 1)     &           &           & -0.034    & -0.034    & -0.033    & -0.027    \\
    Share of Women ($t$ - 1)   &           &           &           & -0.225    & -0.176    & 0.010     \\
    Empl. Rate Women ($t$ - 1) &           &           &           &           & 0.178     & 0.280     \\
    Childcare 0-2 ($t$ - 1)    &           &           &           &           &           & -0.239    \\
    Childcare 3-5 ($t$ - 1)    &           &           &           &           &           & 0.048     \\ 
    \hline & \\[-1.0em]
    \multicolumn{7}{c}{C: Wage Gap at Mean}                                                           \\ 
    \hline & \\[-1.0em]
    Bite x 2018               & -1.609*   & -1.339    & -1.634*   & -1.684*   & -2.139**  & -2.297**  \\
    2018                      & -2.154*** & -3.511*** & -3.163*** & -2.838*   & -7.110**  & -6.581**  \\
    GDP per capita ($t$ - 1)   &           & 0.289     & 0.292     & 0.301     & 0.249     & 0.229     \\
    Pop. Density ($t$ - 1)     &           &           & -0.032    & -0.031    & -0.023    & -0.018    \\
    Share of Women ($t$ - 1)   &           &           &           & 1.006     & 1.289     & 1.662     \\
    Empl. Rate Women ($t$ - 1) &           &           &           &           & 1.014*    & 1.092*    \\
    Childcare 0-2 ($t$ - 1)    &           &           &           &           &           & -0.228    \\
    Childcare 3-5 ($t$ - 1)    &           &           &           &           &           & 0.007     \\ 
    & \\[-1.0em]
    \hline & \\[-1.0em]
    Observations              & 514       & 514       & 514       & 514       & 514       & 514       \\
    Groups                    & 257       & 257       & 257       & 257       & 257       & 257       \\
    \hline & \\[-1.0em]
    \end{tabular}

    }
    Source: SES 2010, 2014 and 2018, own calculations. Note: We only include eligible employees that are not employed in the public service. The subsample is identiﬁed via the variable GG2010, which is provided by the federal statistical oﬃce and which adapts the 2014 and 2018 data to the structure of the 2010 wave (FDZ, 2019).
    \label{tab:main}
    \end{table}

@@

@@

% 25th percentile
% significant, but smaller results.
For the 
% (c6) -18\% compared to the level in 2014 (which was at 18.3%)
% Explain why result at the 25th percentile parts over 30 percent of affected women, compare figure 4
% Additional increases for low payed employees expected (source?)

@@

% (c6) -2.3 ,  relative reduction of 11% 2014 level which was 20.4%)
%  dummy negative -> general reduction, more in treatment group
% Explain why result at the 25th percentile parts over 30 percent of affected women, compare figure 4
Given that the minimum wage directly only affected a small fraction of women working (14.65\% earned less than 8.84€ in 2014), one might ask the question why there is an effect at the 25th percentile at all. There are two relevant explanations for that.

Spillover effect for incomes up to 12.50€ \citep{Dustman21Reallocation} 
\subsection{Robustness, Validity and other fun stuff}

>>>>

\documentclass[12pt,draft,a4paper]{article}
\usepackage[authoryear]{natbib}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{caption}
\usepackage{newtx}
\usepackage[a4paper, left=3cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage[noabbrev]{cleveref}
\captionsetup[table]{labelfont=bf,
                     labelsep=newline,
                     singlelinecheck=false}

@@

    \author{Moritz Trinkner \thanks {sad }}
    \title{Seminar:\\ Quantitative Methoden}
    \date{February 2024}
\end{titlepage}
\maketitle
\thispagestyle{empty}
\newpage
\tableofcontents
\thispagestyle{empty}
\newpage
\setcounter{page}{1}

@@

Table 1 shows the descriptive statistics for all the variables we used in our analysis. We have 40,295 mothers in our sample.
Noticeable values in the table are that only 13.5\% of mothers are working and accordingly the average hours worked per mother is only 5.
The maximum working hours might seem high,
but since only 1\% of employed mothers work more than 70 hours and there are reports of sweatshops in Jordan with up to 100h worked per week \citep{NYT06} we deemed it plausible.
It is worth to note here, that the working hours per week, only looking at working mothers, is 40.2.
This suggests that part time work is almost non-existent in Jordan, in our sample less than 4\% of working mothers work less that 30 hours per week.
This lack of part time work can partially be of an explanation of the low workforce participation of women.

Additionally, 97.6\% of mothers are married and 81.1\% have more than 1 child.
All of this points to the fact that Jordan is a highly conservative and religious country with strict gender roles.
In our sample 51.7\% of first-born kids are boys, being close to the natural sex ratio. Therefore the key identification assumption that gender of first born boy is random holds.

\begin{table}[hbp]
    \caption{T-tests of mothers' characteristics on sex of the first-born child}
    \begin{tabularx}{\textwidth}{lllll}
        \hline
          & \begin{tabular}[c]{@{}l@{}}Mothers with\\ first-born boys\\ (St. err.)\end{tabular} & 
          \begin{tabular}[c]{@{}l@{}}Mothers with\\ first-born girls\\ (St. err.)\end{tabular} & 
          \begin{tabular}[c]{@{}l@{}}Difference\\ (St. err.) \\ \end{tabular} & 
          \begin{tabular}[c]{@{}l@{}}T-test statistic\\ (P-value) \\ \end{tabular} \\ \hline
    Employed              & 0.134           & 0.132        & -0.002   & -0.612       \\
                          & (0.002)         & (0.002)      & (0.003)  & (0.541)      \\
    Hours worked per week & 5.471           & 5.417        & -0.053   & -0.371       \\
                          & (0.100)         & (0.103)      & (0.143)  & (0.710)      \\
    Age                   & 30.643          & 30.721       & 0.078    & 1.344        \\
                          & (0.040)         & (0.042)      & (0.058)  & (0.179)      \\
    Married               & 0.975           & 0.976        & 0.001    & 0.833        \\
                          & (0.001)         & (0.001)      & (0.002)  & (0.405)      \\
    More than 1 child     & 0.805           & 0.819        & 0.014    & 3.617        \\
                          & (0.003)         & (0.003)      & (0.004)  & (0.000)      \\
                          \hline
                        \end{tabularx}

@@

\begin{table}[tbp]% not h om its own
    \caption{First child gender, fertility and marital status of the mother}

@@

    \hline & \\[-1.0em]
    Hours worked per week & & Probability of working &  \\
    \hline & \\[-1.0em]
        First-born boy          & 0.086           & First-born boy           & 0.003           \\
        St. err.                & 0.141           & St. err.                 & 0.003           \\
        \\[-1.0em]
        Baseline Girl:          & 5.378           & Baseline Girl:           & 0.133           \\
        St. err.                & 1.524           & St. err.                 & 0.036           \\
        Percent effect          & 1.596           & Percent effect           & 2.172           \\
        No. obs.                & 40,295          & No. obs.                 & 40,295          \\
    \hline
    \end{tabularx}

@@

\begin{table}[bp]% not h om its own
    \caption{First child gender, fertility and marital status of the mother}

@@

The following section all summarizes the paper “Did the minimum wage reduce the gender wage gap in Germany?” by \citet{CALIENDO22}. It addresses the question if the minimum wage can have an effect on the gender wage gap. 
The reason why this effect might appear is that women are in most countries and regions disproportionally represented among low paid workers. \citep{kahn2015wage}
This means that an increase in the minimum wage would affect women's average wages more than mens, leading to a decrease in the gender wage gap.
The minimum wage is a national wage floor, that is binding for almost all employees. Germany is a special research case, since it was one of the last OECD countries to introduce a Minimum wage. 
When it was introduced in 2015 at 8.50€ it was one of the highest in the EU in terms of purchasing power parity. It affected 10\% of the workforce, and 2/3 of those affected low income workers were women.  
This strong increase makes Germany a great natural experiment to observe effects of the minimum wage, in this case the effect of the gender wage gap. 
%Short preview of results in introduction
%Detailed OECD Data from 2014 First Decile comparison to average
% not a panel study

@@

% Maybe explain ex
The Method used in the paper is a difference-in-difference approach with fixed effects. 
Since there are no legislative differences for the federal wage in difference regions in Germany, the authors use regional differences in Germany to assign Germany to the control or treatment group.
%Why fraction of in relation to women and not the whole workforce
%Not dependent on gender wage gap neither on fraction of men working under the minimum wage
%"We generate the gender wage gap at the regional level as the difference between men’s and women’s wages divided by men’s wages."
This is followed after \citet{Card1992} who did the same in an analysis of the minimum wage in the United States. 
They  use labor market regions to split the country in 257 different regions. 
The fraction of women that worked below the minimum wage in 2014 in relation to all employed women is used to see how 
strong %better word 
the introduction of the minimum wage affects the employed women in the region. 
\Citeauthor{CALIENDO22} call this the bite, which can be visualized by how deep the minimum wage bites into the wage distribution.
%median bite level of 17.15%.
The fraction for all regions is used to separate them at the median, so that regions with more than 17.15\% of women affected by the minimum wage are in the treatment group and regions below the control group.

@@

\subsection{Main Result}
% • What is the main result? And what is the implication of this finding?
% Summarize and synthesize the results, provide relevant background information

@@

\begin{table}[htbp] %Main Results
    \caption{Regressions of wage gaps at 10th percentile, 25th percentile and mean.}
    %\centering
    %\resizebox{\textwidth}{!}{%
    \begin{adjustbox}{width=\textwidth}
    \begin{tabular}{l@{\hskip 10ex}llllll}
    \hline & \\[-1.0em]
                              & (1)       & (2)       & (3)       & (4)       & (5)       & (6)       \\ 
    \hline & \\[-1.0em]
    \multicolumn{7}{c}{A: Wage Gap at p10}                                                            \\
    \hline & \\[-1.0em]
    Bite x 2018               & -5.228*** & -5.085*** & -4.870*** & -4.735*** & -4.550*** & -4.550*** \\
    2018                      & -3.568*** & -4.286*** & -4.539*** & -5.419*** & -3.684    & -3.754    \\
    GDP per capita ($t$ - 1)   &           & 0.153     & 0.152     & 0.126     & 0.147     & 0.158     \\
    Pop. Density ($t$ - 1)     &           &           & 0.024     & 0.020     & 0.017     & 0.015     \\
    Share of Women ($t$ - 1)   &           &           &           & -2.722    & -2.837    & -2.610    \\
    Empl. Rate Women ($t$ - 1) &           &           &           &           & -0.412    & -0.457    \\
    Childcare 0-2 ($t$ - 1)    &           &           &           &           &           & 0.044     \\
    Childcare 3-5 ($t$ - 1)    &           &           &           &           &           & -0.061    \\ 
    \hline & \\[-1.0em]
    \multicolumn{7}{c}{B: Wage Gap at p25}                                                            \\ 
    \hline & \\[-1.0em]
    Bite x 2018               & -3.352*** & -2.814*** & -3.122*** & -3.111*** & -3.191*** & -3.336*** \\
    2018                      & -1.454**  & -4.154*** & -3.792*** & -3.865**  & -4.612    & -4.079    \\
    GDP per capita ($t$ - 1)   &           & 0.576**   & 0.578**   & 0.576**   & 0.567**   & 0.541**   \\
    Pop. Density ($t$ - 1)     &           &           & -0.034    & -0.034    & -0.033    & -0.027    \\
    Share of Women ($t$ - 1)   &           &           &           & -0.225    & -0.176    & 0.010     \\
    Empl. Rate Women ($t$ - 1) &           &           &           &           & 0.178     & 0.280     \\
    Childcare 0-2 ($t$ - 1)    &           &           &           &           &           & -0.239    \\
    Childcare 3-5 ($t$ - 1)    &           &           &           &           &           & 0.048     \\ 
    \hline & \\[-1.0em]
    \multicolumn{7}{c}{C: Wage Gap at Mean}                                                           \\ 
    \hline & \\[-1.0em]
    Bite x 2018               & -1.609*   & -1.339    & -1.634*   & -1.684*   & -2.139**  & -2.297**  \\
    2018                      & -2.154*** & -3.511*** & -3.163*** & -2.838*   & -7.110**  & -6.581**  \\
    GDP per capita ($t$ - 1)   &           & 0.289     & 0.292     & 0.301     & 0.249     & 0.229     \\
    Pop. Density ($t$ - 1)     &           &           & -0.032    & -0.031    & -0.023    & -0.018    \\
    Share of Women ($t$ - 1)   &           &           &           & 1.006     & 1.289     & 1.662     \\
    Empl. Rate Women ($t$ - 1) &           &           &           &           & 1.014*    & 1.092*    \\
    Childcare 0-2 ($t$ - 1)    &           &           &           &           &           & -0.228    \\
    Childcare 3-5 ($t$ - 1)    &           &           &           &           &           & 0.007     \\ 
    & \\[-1.0em]
    \hline & \\[-1.0em]
    Observations              & 514       & 514       & 514       & 514       & 514       & 514       \\
    Groups                    & 257       & 257       & 257       & 257       & 257       & 257       \\
    \hline & \\[-1.0em]
    \end{tabular}
    \end{adjustbox}
    %}

    Note: which adapts the 2014 and 2018 data to the structure of the 2010 wave (FDZ, 2019).
    \label{tab:main}

    \end{table}

@@

% 25th percentile
% significant, but smaller results.
For the 25th percentile
% (c6) -18\% compared to the level in 2014 (which was at 18.3%)
% Explain why result at the 25th percentile parts over 30 percent of affected women, compare figure 4
% Additional increases for low payed employees expected (source?)

@@

% (c6) -2.3 ,  relative reduction of 11% 2014 level which was 20.4%)
%  dummy negative -> general reduction, more in treatment group
% Explain why result at the 25th percentile parts over 30 percent of affected women, compare figure 4
Given that the minimum wage directly only affected a small fraction of women working (14.65\% earned less than 8.84€ in 2014), one might ask the question why there is an effect at the 25th percentile at all.
%Rewrite that sentence 
There are two relevant explanations for that.
First thing to note is, that while the cutoff for the enrollment in the treatment group is only 17.5\%,
some of the regions (and almost all of the regions in eastern Germany)
had more than 25\% of women affected. This leads to the minimum wage directly affecting their wages even at and above the 25th percentile of the wage distribution, and thus leading to a reduction in the gender wage gap there.
Another reason is the spillover effect, meaning that an increase in the minimum wage not only raises wages below the wage floor, but also wages slightly above it.
In a paper analyzing these spillover effects in the UK, \citet{Stewart12Wage} cites multiple reasons for this, like the continuity of existing wage differentials or higher reservation wages for certain jobs and sees the minimum wage as a general increase in the cost of low-skilled labor.
For Germany, \citet{Dustman21Reallocation} find that spillover effects due to the minimum wage introduction effect initial wages up to 12.50€, which is just slightly below the median for women in 2014 of 12.54€ \citep[see][Table 1]{CALIENDO22}. 
%Spillover effect for incomes up to 12.50€ \citep{Dustman21Reallocation} 
\subsection{Robustness, Validity and other fun stuff}
